\chapter{Internal Definitions}

\appnpara The following definitions shall apply for the purposes of interpreting the Standing Orders.
\begin{description}
\item[Academic Affairs Representative] the Officer so described in S.O. 3.8.
\item[Access and Equal Opportunities Representative] the Officer so described in S.O. 3.7.
\item[Access Sub-Committee] the Sub-Committee so described in S.O. 2.6.3.
\item[Admiral of the Fleet] the Sports and Recreation Representative
\item[Affiliated Organisations Representative] the Officer so described in S.O. 3.11.
\item[Arts and Culture Representative] the Officer so described in S.O. 3.11.
\item[Associate Member] an associate member of the JCR Committee in accordance with S.O. 2.3.4.
\item[BME Representative] the Ethnic Minority Students’ Representative.
\item[Charities Representative] the Officer so described in S.O. 3.13.
\item[Committee Chair] the Officer designated as such in S.O. 2.3.4.
\item[Disabled Students’ Representative] the Officer so described in S.O. 3.14.
\item[Entertainment Representative] one of the Officers so describes in S.O. 3.10.
\item[Environment Representative] the Officer so described in S.O. 3.15.
\item[Equality Sub-Committee] the Sub-Committee so described in S.O. 2.6.4.
\item[Ethnic Minority Students' Representative] the Officer so described in S.O. 3.16.
\item[Events Photographer] the Officer so described in S.O. 3.24.
\item[Food and Bar Representative] the Officer so described in S.O. 3.17.
\item[Freshers' Week] noughth Week of Michaelmas Term
\item[Gender Equality Representative] the Officer so described in S.O. 3.18.
\item[International Students' Representative] the Officer so described in S.O. 3.19.
\item[LGBTQIA+ Representative] the Officer so described in S.O. 3.20.
\item[Mandating Resolution] a resolution of a motion designated so by S.O. 6.9.1.6.
\item[Memorandum] any understanding between the JCR and another party constituting best practice 
\item[Office-Holder] an Officer or a member of a Standing Committee or Sub-Committee
\item[Publications and Social Media Officer] the Officer so described in S.O. 3.21.
\item[Record] a document designated as a record by S.O. 11.
\item[Room Ballot] the room ballot administered in accordance with S.O. 7.
\item[Sports and Recreation Representative] the Officer so described in S.O. 3.22.
\item[Subject President] one of the Officers so described in S.O. 3.25.
\item[Suspended Students’ Representative] the Officer so described in S.O. 3.23.
\item[Welfare Officer] one of the Officers so described in S.O. 3.9.
\item[Welfare Sub-Committee] the Sub-Committee so described in S.O. 2.6.5.
\end{description}